---
title: "SLU DSS Spring 2020"
subtitle: "Statistics in R: Session #2"
author: "Cort W. Rudolph"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: ioslides_presentation
widescreen: true
smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = FALSE,
  message = FALSE,
  warning = FALSE
  )
options(scipen = 99, digits = 3)
```

# Statistics in R|Session #2: Modeling Continuous Predictors

## Needed packages
- Let's now load all the packages needed for this week. 
- `ipak` codeblock available in the `RMarkdown` file:

<!-- To install & load the required packages for this lecture, run this codeblock -->
```{r echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
# ipak <- function(pkg){
#     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
#     if (length(new.pkg))
#         install.packages(new.pkg, dependencies = TRUE)
#     sapply(pkg, require, character.only = TRUE)
# }
# 
# # usage
# packages <- c("tidyverse",
#               "skimr",
#               "gridExtra",
#               "kableExtra",
#               "psych",
#               "broom",
#               "magrittr",
#               "ggeffects")
# ipak(packages)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(psych)
library(gridExtra)
library(kableExtra)
library(broom)
library(magrittr)
library(ggeffects)
```

## Modeling Continuous Predictors
- Professor rating data
- Let's load the data, `select` only a subset of the variables:

```{r eval=TRUE}
load(url("http://www.openintro.org/stat/data/evals.RData"))
evals <- evals %>%
  select(score, bty_avg)
```

## Modeling Continuous Predictors
- Correlation, $r_{xy}$, between `score` and `bty_avg`

```{r}
evals %>% 
  cor()
```

## Modeling Continuous Predictors
- 95% confidence intervals around correlation

```{r}
evals %$%
  cor.test(score, bty_avg) %>% 
  broom::tidy() %>%
  select(conf.low, estimate,conf.high) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

## Modeling Continuous Predictors

```{r, eval = TRUE, warning=FALSE, fig.cap="Instructor Evaluation Scores at UT Austin"}
ggplot(evals, aes(x = bty_avg, y = score)) + 
       geom_jitter() +  
       labs(x = "Beauty Score", y = "Teaching Score")
```

## Modeling Continuous Predictors
- Let's improve this scatterplot by adding a "regression line." 
- This is done by adding a new layer to the `ggplot` code that created the original figure: `+ geom_smooth(method = "lm", se = FALSE)`.
- A regression line is a "best fitting" line in that of all possible lines you could draw on this plot; it is "best" in terms of  the "least squares" criteria.


## Modeling Continuous Predictors
```{r, eval=TRUE, warning=FALSE, fig.cap="Regression Line"}
ggplot(evals, aes(x = bty_avg, y = score)) + 
  geom_jitter() +
  labs(x = "Beauty Score", y = "Teaching Score") + 
  geom_smooth(method = "lm", se = FALSE)
```

## Modeling Continuous Predictors
- Here, the regression line is a nice visual summary of the relationship between these two numerical variables.
- The positive slope of the blue line is consistent with our observed correlation coefficient of `r cor(evals$score, evals$bty_avg) %>% round(3)` suggesting that there is a positive relationship between `score` and `bty_avg`. 

## Modeling Continuous Predictors
- Let's build a statistical model to represent this relationship

```{r}
evals %$%
  lm(score ~ bty_avg) %>%
  broom::tidy(conf.int=TRUE) %>%
  select(estimate, std.error, conf.low, conf.high) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

## Modeling Continuous Predictors
- Let's estimate $R^2$ from this model

```{r}
evals %$%
  lm(score ~ bty_avg) %>% 
  summary() %$% 
  r.squared %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

## Modeling Continuous Predictors
Recall that our correlation was `r cor(evals$score, evals$bty_avg) %>% round(3)`

```{r}
evals %$% 
  cor(score, bty_avg)^2
```

## Modeling Continuous Predictors
- Finally, plot predicted values from this model:
```{r}
evals %$%
  lm(score ~ bty_avg) %>%
  ggeffects::ggpredict() %>%
  plot()
```



# Your Turn!

## CDC Data
The Behavioral Risk Factor Surveillance System (BRFSS) is an annual telephone survey of 350,000 people in the United States conducted by the CDC

Here, we will focus on a random sample of 20,000 people from the BRFSS survey conducted in 2000. While there are over 200 variables in this data set, for now, we will work with a small subset (i.e., just respondents height & weight):

First, we load the data:
```{r}
source("http://www.openintro.org/stat/data/cdc.R")
```

Then, select a random sample of $n = 100$ respondents height & weight
```{r}
set.seed(1818)
cdc_subset<-cdc %>%
  sample_n(100) %>%
  select(height, weight)
```

## What is the correlation between `height` and `weight`?
```{r}
cdc_subset %>% 
  cor()
```

## What is 95% confidence interval around the correlation between`height` and `weight`?
```{r}
cdc_subset %$%
  cor.test(height, weight) %>% 
  broom::tidy() %>%
  select(conf.low, estimate,conf.high) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

## Create a scatterplot of the relationship between `height` and `weight`.
```{r}
ggplot(cdc_subset, aes(x = height, y = weight)) +
       geom_jitter() +
       labs(x = "Height", y = "Weight") +  
       geom_smooth(method = "lm", se = FALSE)
```

## Regress`weight` onto `height`, and create a table of parameter estimates.

```{r}
cdc_subset %$%
  lm(weight ~ height) %>%
  broom::tidy(conf.int=TRUE) %>%
  select(estimate, std.error, conf.low, conf.high) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

## What percentage of variance in `weight` can be explained by `height`?

```{r}
cdc_subset %$%
  lm(weight ~ height) %>% 
  summary() %$% 
  r.squared %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

## Plot the predicted relationship between `height` and `weight`
```{r}
cdc_subset %$% lm(weight ~ height) %>% 
  ggeffects::ggpredict() %>%
  plot()
```

